<!-- 
This file is part of a gitbook that should be cited as: 

XX

This tutorial has been supported by the XX project
-->

# Download

The Copernicus Marine Toolbox is written in Python, but it can be used in R through the `reticulate` package. For more information on how to use the Copernicus Marine Toolbox in R, see the [official documentation](https://help.marine.copernicus.eu/en/articles/8638253-how-to-download-data-via-the-copernicus-marine-toolbox-in-r).

## Installation

First, we need to install the `reticulate` package:

```{r, eval=F}
install.packages("reticulate") 
```

Then, the package can be loaded:

```{r, eval=F}
library("reticulate")
```

Now we create a dedicated Python virtual environment for the Copernicus Marine Toolbox. Using a separate environment ensures a clean, isolated space where the package and its dependencies won’t interfere with other projects. The virtual environment has been created in 

```{r, eval=F}
virtualenv_create(envname = "CopernicusMarine")

virtualenv_install("CopernicusMarine", packages = c("copernicusmarine"))

use_virtualenv("CopernicusMarine", required = TRUE)
```



